---
title: "Modelling public road accidents"
author: "Katarzyna Kobalczyk"
date: "22/08/2021"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(bnlearn)
library(scales)
library(stagedtrees)

theme_set(theme_minimal())
```

# Base model

## Data Preprocessing

Variables included in the base regression model:
- `sex`
- `age` - age at which the respondent passed the practical driving test
- `experience` -  number of years the respondent has been driving
- `exposure` - defined as the annualised number of miles driven + 10 x the annualised number of days on which the driver has driven
- `accident liability`- annualised accidents

```{r}
df <- read.table("data/deq1.tab", sep = "\t", header = TRUE)

impute_missing <- function(x) {ifelse(x < 0, NA, x)}

df <- df %>% 
  select(sex, age = Age, miles = V00138, freq = V00160, acc1, nlacc1, 
         style1 = ds1, style2 = ds2, style3 = ds3,
         Hazardscore, violations, errors, aggressiveviolations,
         inexperience, slips) %>%
  apply(2, impute_missing) %>%
  as.data.frame() %>%
  mutate(sex = fct_recode(factor(sex), F = "1", M = "2"),
         acc_inv = factor(case_when(
           is.na(nlacc1)  ~ "NA", 
           nlacc1 >= 1 ~ "1+",
           TRUE ~ "0")),
         miles = miles * 2,
         freq_lab = fct_recode(factor(freq),
                              "Everyday" = "1",
                              "4-6 days per week" = "2",
                              "1-3 days per week" = "3",
                              "About a fortnight" = "4",
                              "About once a month" = "5",
                              "Less than once a month" = "6",
                              "Never" = "7") # annualised milleage
  )

df
```
## EDA

### Sex and age

```{r}
df %>% ggplot(aes(x = sex, fill = sex)) + 
  geom_bar()

df %>% ggplot(aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "white", colour = "black")

df %>% ggplot(aes(x = sex, y = age, colour = sex)) +
  geom_boxplot()
```
### Milleage - annualised and Frequency

```{r}
summary(df$miles)

n_obs <- sum(!is.na(df$miles))
  
df %>% 
  filter(!is.na(miles)) %>%
  mutate(bins = cut(miles, breaks = c(-Inf, 1000, 5000, 10000, 15000, Inf))) %>%
  group_by(bins) %>%
  count() %>%
  ggplot(aes(x = bins, y = n/n_obs * 100)) + 
  geom_bar(stat = "identity") + 
  ylab("% of respondents") + 
  xlab("Annualised milleage")

df %>% 
  filter(!is.na(freq)) %>%
  ggplot(aes(x = freq_lab)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
```{r}
df %>%
  filter(!is.na(miles) & !is.na(freq)) %>%
  filter(!(abs(miles - median(miles) > 2*sd(miles)))) %>%
  ggplot(aes(x = freq_lab, y = miles)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  xlab("Frequency of driving in first 6 months") + 
  ylab("Annualised milleage") + 
  ggtitle("Milleage vs. frequency", subtitle = "(extreme values removed)") + 
  scale_y_continuous(trans = "log1p")
```


### Predictors vs. accidents

```{r}
df %>%
  filter(!is.na(acc_inv)) %>%
  ggplot(aes(x = factor(acc_inv), y = age)) +
  geom_boxplot() + 
  scale_y_continuous(trans = "log") + 
  xlab("Involved involvement") + 
  ylab("Age")

df %>%
  filter(!is.na(acc_inv) & !is.na(miles)) %>%
  ggplot(aes(x = factor(acc_inv), y = miles)) +
  geom_boxplot() + 
  scale_y_continuous(trans = "log1p") + 
  xlab("Involved involvement") + 
  ylab("Annualised milleage")

table(df %>% select(freq_lab, acc_inv))

df %>%
  filter(!is.na(acc_inv) & !is.na(freq_lab)) %>%
  group_by(freq_lab, acc_inv) %>%
  count() %>%
  ggplot(aes(x = freq_lab, y = n, fill = acc_inv)) +
  geom_bar(stat = "identity", position = "fill") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

by_ages <- df %>% 
  mutate(age_group = cut(df$age, 
                         breaks = c(-Inf, 18, 20, 25, Inf))) %>%
  select(age_group, acc_inv) %>%
  table()

by_ages / rowSums(by_ages)

df %>% 
  mutate(age_group = as.integer(age)) %>%
  filter(!is.na(acc_inv)) %>%
  ggplot(aes(x = age_group, group = acc_inv, fill = acc_inv)) +
  geom_bar() + 
  #scale_y_continuous(trans = "log1p") + 
  xlab("age")

df %>% 
  mutate(age_group = as.integer(age)) %>%
  filter(!is.na(acc_inv)) %>%
  ggplot(aes(x = age_group, group = acc_inv, fill = acc_inv)) +
  geom_bar(position = "fill") + 
  xlab("age")
```

```{r}
df %>%
  filter(!is.na(ac_acc_inv)) %>%
  mutate(age_group = cut(age, breaks = c(-Inf, 18, 20, 25, 30, 40, Inf))) %>%
  group_by(age_group, sex, ac_acc_inv) %>%
  count() %>%
  ggplot(aes(x = age_group, y = n, fill = ac_acc_inv)) +
  geom_bar(stat = "identity", position = "fill") + 
  facet_wrap(~sex)

table(df %>% select(ac_acc_inv, sex)) / rowSums(table(df %>% select(ac_acc_inv, sex)))
```

## Bayesian Network

```{r}
unique(cut_number(df$age, 4))

df <- df %>% 
  mutate(age_group = cut(age, breaks = c(16, 18, 20, 25, 80)))

mod_df <- df %>%
  select(age_group, sex, freq_lab, acc_inv) %>%
  drop_na(freq_lab) %>%
  mutate(acc_inv = if_else(freq_lab == "Never", "0", as.character(acc_inv)),
         acc_inv = factor(acc_inv)) %>%
  filter(acc_inv != "NA") %>%
  mutate(acc_inv = droplevels(acc_inv, "NA"))
  
#table(mod_df)
```

```{r}
bn <- bnlearn::hc(mod_df, score = "aic", blacklist = data.frame(from = c("acc_inv"), to = c("sex")))

plot(bn)
```

```{r}
bn_tree <- as_sevt(bn.fit(bn, data = mod_df), order  = c("age_group", "sex", "freq_lab", "acc_inv"))
bn_tree <- sevt_fit(bn_tree, data = mod_df)
plot(bn_tree) #,col = function(s) hue_pal()(length(s)))

barplot(bn_tree, var = "acc_inv")#, col = hue_pal()(4))
BIC(bn_tree)
```
```{r}
# reduce the number of freq categories
table(mod_df %>% select(freq_lab, acc_inv)) 
table(mod_df %>% select(freq_lab, acc_inv)) / rowSums(table(mod_df %>% select(freq_lab, acc_inv)))

mod_df2 <- df %>%
  select(age_group, sex, freq, freq_lab, acc_inv) %>%
  drop_na(freq_lab) %>%
  mutate(acc_inv = if_else(freq_lab == "Never", "0", as.character(acc_inv)),
         acc_inv = factor(acc_inv)) %>%
  filter(acc_inv != "NA") %>%
  mutate(acc_inv = droplevels(acc_inv, "NA")) %>%
  mutate(freq2 = case_when(
    freq == 1 ~ 1,
    freq == 2 ~ 2,
    freq == 3 ~ 3,
    freq %in% c(4, 5) ~ 4,
    freq %in% c(6, 7) ~ 5
  ),
  freq2 = factor(freq2)) %>%
  select(age_group, freq2, sex, acc_inv) %>%
  na.omit()

table(mod_df2)

bn2 <- bnlearn::hc(mod_df2, score = "aic" , blacklist = data.frame(from = c("acc_inv"), to = c("sex")))
plot(bn2)
```

```{r}
bn_tree2 <- as_sevt(bn.fit(bn2, data = mod_df2), order  = c("age_group", "sex", "freq2", "acc_inv"))
bn_tree2 <- sevt_fit(bn_tree2, data = mod_df2)
plot(bn_tree2)

barplot(bn_tree2, var = "acc_inv")

cbind(BIC(bn_tree2, bn_tree), AIC(bn_tree2, bn_tree))
```

```{r}
order <- c("age_group", "sex", "freq2", "acc_inv")

mod1 <- mod_df2 %>%
  indep(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_hc()

mod2 <- mod_df2 %>%
  full(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_hc() 

mod3 <- mod_df2 %>%
  full(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_bj(distance = "kullback", thr = 0.035)

cbind(BIC(bn_tree2, mod1, mod2, mod3), AIC = AIC(bn_tree2, mod1, mod2, mod3)$AIC)
```

```{r}
order <- c("freq2", "age_group", "sex", "acc_inv")

mod1a <- mod_df2 %>%
  indep(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_hc()

mod2a <- mod_df2 %>%
  full(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_hc() 

mod3a <- mod_df2 %>%
  full(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_bj(distance = "kullback", thr = 0.035)

cbind(BIC(mod1a, mod2a, mod3a), AIC = AIC(mod1a, mod2a, mod3a)$AIC)
```


```{r}
# Best BIC model the same as best 
best_mod <- search_best(mod_df2, join_unobserved = FALSE, alg = stages_hc)
plot_mod(best_mod)  
```


```{r}
plot_mod <- function(mod) {
  plot(mod)
  barplot(mod, var = "acc_inv")
  plot(ceg(mod))
}
```



```{r}
plot(mod1)
barplot(mod1, var = "acc_inv")
plot(ceg(mod1))
```

```{r}
plot(mod1a)
barplot(mod1a, var = "acc_inv")
plot(ceg(mod1a))
```
### Combining sex and age

```{r}
mod_df3 <- mod_df2 %>%
  mutate(age_sex = factor(paste0(age_group, "-",sex)))

new_bn <- bnlearn::hc(mod_df3 %>% select(age_sex, freq2, acc_inv), score = "aic")
plot(new_bn)
```
```{r}
order <- c("age_sex", "freq2", "acc_inv")

mod_new <- mod_df3 %>%
  indep(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_hc()
```

```{r}
plot(mod_new)
barplot(mod_new, var = "acc_inv")
plot(ceg(mod_new))
```

# Missingness of the response

```{r}
mod_mis_df <- df %>%
  select(age_group, sex, freq, freq_lab, acc_inv) %>%
  drop_na(freq_lab) %>%
  mutate(acc_inv = if_else(freq_lab == "Never", "0", as.character(acc_inv)),
         acc_inv = factor(acc_inv),
         is_missing = factor(if_else(acc_inv == "NA", "yes", "no"))) %>%
  mutate(freq2 = case_when(
    freq == 1 ~ 1,
    freq == 2 ~ 2,
    freq == 3 ~ 3,
    freq %in% c(4, 5) ~ 4,
    freq %in% c(6, 7) ~ 5
  ),
  freq2 = factor(freq2)) %>%
  select(age_group, sex, freq2, is_missing)

order <- c("age_group", "sex", "freq2", "is_missing")

bn_mis <- bnlearn::hc(mod_mis_df, score = "bic")
plot(bn_mis)
```

## Hazard Perception

```{r}
df_full <- read.table("data/unified_database_2014.tab", sep = "\t", header = TRUE)

df_full <- df_full %>% 
  select(ID, sex = Sex_all, cohort = Cohort_all, age = Age_all, miles = t1_a2, 
         freq = t1_a1, nlacc1 = t1_nlacc_orig, hp =  t0_takeHP) %>%
  apply(2, impute_missing) %>%
  as.data.frame() %>%
  mutate(sex = fct_recode(factor(sex), F = "0", M = "1"),
         acc_inv = factor(case_when(
           is.na(nlacc1)  ~ "NA", 
           nlacc1 >= 1 ~ "1+",
           TRUE ~ "0")),
         miles = miles * 2,
         hp = factor(hp))

df_full
```
```{r}
table(df_full[,c("cohort", "hp")])
```


```{r}
mod_hp_df <- df_full %>%
  mutate(age_group = cut(age, breaks = c(16, 18, 20, 25, 80)),
         age_sex = paste0(age_group, "-", sex)) %>%
  drop_na(freq) %>%
  mutate(acc_inv = if_else(freq == "7", "0", as.character(acc_inv)),
         acc_inv = factor(acc_inv)) %>%
  filter(acc_inv != "NA") %>%
  mutate(acc_inv = droplevels(acc_inv, "NA")) %>%
  mutate(freq2 = case_when(
    freq == 1 ~ 1,
    freq == 2 ~ 2,
    freq == 3 ~ 3,
    freq %in% c(4, 5) ~ 4,
    freq %in% c(6, 7) ~ 5
  ),
  freq2 = factor(freq2)) %>%
  select(hp, age_group, age_sex, freq2, sex, acc_inv) %>%
  na.omit()

mod_hp_df
```
```{r}
order <- c("freq2", "hp", "acc_inv")

mod1_hp <- mod_hp_df %>%
  indep(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_hc()

plot(mod1_hp)
barplot(mod1_hp, var = "acc_inv")
plot(ceg(mod1_hp))
```


```{r}
order <- c("freq2", "sex", "hp",  "acc_inv")

mod1_hp <- mod_hp_df %>%
  indep(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_hc()

plot(mod1_hp)
barplot(mod1_hp, var = "acc_inv")
plot(ceg(mod1_hp))
```
```{r}
order <- c("freq2", "age_group", "hp",  "acc_inv")

mod1_hp <- mod_hp_df %>%
  indep(order = order, join_unobserved = FALSE, lambda = 0) %>%
  stages_hc()

plot(mod1_hp)
barplot(mod1_hp, var = "acc_inv")
plot(ceg(mod1_hp))
```
```{r}
order <- c("age_sex", "hp", "acc_inv")

mod1_hp <- mod_hp_df %>%
  indep(order = order, join_unobserved = FALSE, lambda = 0.5) %>%
  stages_hc()

plot(mod1_hp)
barplot(mod1_hp, var = "acc_inv")
plot(ceg(mod1_hp))
```
```{r}
df_full %>%
  drop_na(nlacc1) %>%
  group_by(cohort) %>%
  summarise(acc_pp = sum(nlacc1)/n())
```

