---
title: "CEG - experiments"
author: "Katarzyna Kobalczyk"
date: "19/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(stagedtrees)
library(scales)
library(bnlearn)
```

# Usage of the stagedtrees package

## Learning the stage structure for a dataset

```{r}
data(Titanic)
str(Titanic)
```

```{r}
m_full <- full(Titanic)
m_indep <- indep(Titanic)
m_full
m_indep
plot(m_full, col = function(s) hue_pal()(length(s)))  # option to use different colour palettes
plot(m_indep)
```

```{r}
mod1 <- stages_hc(m_indep)
mod2 <- stages_bj(m_full, thr = 0.1)

plot(mod1)
plot(mod2)
```

```{r}
# models can be refined

mod3 <- stages_hc(mod2)

plot(mod3)
```
```{r}
# comparison of the models

compare_stages(mod1, mod3, plot = TRUE)

cbind(AIC(mod1, mod2, mod3), BIC = BIC(mod1, mod2, mod3)$BIC)
```

````{r}
plot(mod1)
barplot(mod1, "Survived", horiz = TRUE)
```
## Bayesian networks as staged tress

```{r}
titanic.df <- as.data.frame(Titanic)
titanic.df <- titanic.df[rep(row.names(titanic.df), titanic.df$Freq), 1:4]

mod_bn <- bnlearn::hc(titanic.df)
plot(mod_bn)

values = list()
for (c in colnames(titanic.df)) {
  values[[c]] <- unique(titanic.df[,c])
}
bn_tree <- sevt_fit(bn_tree, data = titanic.df, lambda = 0.5)
plot(bn_tree)
```

```{r}
mod4 <- stages_hc(bn_tree, score = function(x) -AIC(x, k = 100))
plot(mod4)
plot(ceg(mod4))
```
```{r}
subtree_crew <- subtree(mod4, c(Class = "Crew"))
subtree_crew
plot(subtree_crew, cex_label_nodes = 1.5, pch = 1)
summary(subtree_crew)
```
```{r}
get_path(mod4, var = "Survived", stage = "5")
```

```{r}
prob(mod4, c(Survived = "Yes"))
prob(mod4, c(Survived = "Yes", Age = "Adult")) / prob(mod4, c(Age = "Adult"))
prob(mod4, c(Survived = "Yes", Age = "Child")) / prob(mod4, c(Age = "Child"))
```

# A dataset analysis using stagedtrees

```{r}
library(magrittr)
data("PhDArticles")
str(PhDArticles)
```

```{r}
bn <- bnlearn::hc(PhDArticles)
plot(bn)

bn_tree <- as_sevt(bn.fit(bn, data = as.data.frame(PhDArticles)), order  = c("Gender", "Kids", "Married", "Articles"))
plot(bn_tree)
```
- Number of Articles of a student is independent of the Gender, Kids and Married
- Given the number of publications of the mentor the number of publications of a student is conditionally independent of the prestige of the university.

We investigate the strength of independence between the Articles and Gender.

```{r}
order <- c("Gender", "Kids", "Married", "Articles")


mod1 <- PhDArticles %>%
  select(all_of(order)) %>%
  indep(order = order) %>%
  stages_hc() %>%
  join_unobserved(name = "NA")

mod2 <- PhDArticles %>%
  select(all_of(order)) %>%
  full(order = order) %>%
  stages_hc() %>%
  join_unobserved(name = "NA")

plot(mod1)
plot(mod2)
```

```{r}
compare_stages(mod1, mod2, plot = TRUE)
```

```{r}
par(mfrow = c(1,2))
plot(mod2)
barplot(mod2, var = "Articles", legend.text = TRUE)
```
Not married women without kids and married women with kids have the lowest proabability of hugh number of articles.
The population with highest probability of high number of articles are males with no kids.

```{r}
get_path(mod2, var = "Articles", "4")
```

We can test if mod1 is sufficient to describe the data against the more complex model mod2 using the LRT.
```{r}
lr_test(mod1, mod2)
```

The small p-value confirms that the asymetric structure described by mod2 is indeed supported by the data.

Finally a staged tree over all the variables is built.
```{r}
mod_all <- PhDArticles %>%
  full(order = c("Prestige", "Mentor", order)) %>%
  join_unobserved() %>%
  stages_bj(thr = 0.5) %>%
  stndnaming()

par(mfrow = c(1,2))
plot(mod_all, cex_label_edges = 0.9)
barplot(mod_all, var = "Articles", legend.text = TRUE)
```

Stage with highest probability of a large number of articles - stage 3 includes the following paths:

```{r}
get_path(mod_all, "Articles", "3")
```
This population only includes students whose mentors have high number of articles. Among them are males from high prestige universities with no kids,
married males and females from low prestige universities and married females with no kids and high prestige universities.

```{r}
plot(ceg(mod_all))
```

